---
title: "hw11_642"
author: "Xuchen Wang"
date: "April 12, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("/Users/xuchenwang/BernBeta.R")
```

#### Problem 1.
#### a) What are the inputs and outputs for BernBeta?  Explain what they mean.
Inputs: priorShape - vector of parameter values for the prior beta distribution.
        dataVec - vector of 1's and 0's: the vector of data for Bernoulli distribution
        credMass - the probability mass of the HDI.
Outputs: postShape - vector of parameter values for the posterior beta distribution.
         Graphics: Creates a three-panel graph of prior, likelihood, and posterior with          highest posterior density interval.
         Example of use:
         postShape = BernBeta( priorShape=c(1,1) , dataVec=c(1,0,0,1,1) )
         You will need to "source" this function before using it, so R knows
         that the function exists and how it is defined.

#### b) Start with a prior of beta(4,4).  Use the BernBeta function to flip the coin once and get a head.  What is the posterior?  Show all plots and explain results
```{r}
poster <- BernBeta(c(4,4),c(1))
poster
```

The posterior is beta(5,4). The results show that after one trial and it gets a head, the distribution of theta move from symmetric to right bias, which means the belief of theta change due to the collected data.

#### c) Use the posterior from the previous flip as the prior for the next flip.  (You can use the result from the last run - "post" - instead of c(5,4) as the input the BernBeta.) Flip the coin again and get a head again.  What is the new posterior?
```{r}
poster2 <- BernBeta(poster,c(1))
poster2
```

The new posterior is beta(6,4)

#### d) Using the posterior from the last flip, flip again and get tails.  What is the new posterior?
```{r}
poster3 <- BernBeta(poster2,c(0))
poster3
```

The new posterior is beta(6,5)

#### e) Start over at b) and do three flips, but in the order T, H, H.  Show the posterior after these 3 flips. How does the posterior change?  Does order of the results matter?
```{r}
poster4 <- BernBeta(c(4,4),c(0,1,1))
poster4
```

The posterior is beta(6,5). It does not change and thus, the order of the results does not matter.

#### f) Start over at b) and do three flips, but in the order T, T, H.  Show the posterior after these 3 flips. How does the posterior change?
```{r}
poster4 <- BernBeta(c(4,4),c(0,0,1))
poster4
```

The posterior is beta(5,6)

#### Problem 2 You have a coin that you know came from a magic shop.  You know it is biased but you are not sure which way, towards heads or tails.

#### a) Find a beta that reflects your beliefs
Beta(1,1)

#### b) Use BernBeta to flip the coin 5 times and get T T H T T.  What is your posterior?  Show all plots and explain your results. 
```{r}
poster_1 <- BernBeta(c(1,1),c(0,0,1,0,0))
poster_1 
```
The posterior is Beta(2,5). The results show that after five trials and it gets T T H T T, the distribution of theta moves from symmetric to right bias since there is more Ts, which means the belief of theta change due to the collected data.

#### c) What will the posterior look like after 100 flips if you get 10 heads and 90 tails? (No need to use the BernBeta function.  Calculate the answer by hand.)
The posterior is Beta(11,91). Since the prior is Beta(1,1) and we get 10 heads and 90 tails, the posterior should be Beta(1+10,1+90)=Beta(11,91).

#### d) Use R to plot the answer to c). 
```{r}
BernBeta(c(1,1),c(rep(1,10),rep(0,90)))
```










