---
title: "HW1_642"
author: "Xuchen Wang"
date: "January 27, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Chapter 2 Problem 10
#### (a)
```{r}
library(MASS)
data(Boston)
?Boston
dim(Boston) 
```

506 rows and 14 columns. Rows represent observations. Columns represent predictors.

#### (b)
```{r}
attach(Boston)
pairs(Boston)       # plot all pairwise scatterplots
par(mfrow=c(2,2))
plot(zn, rm)        
plot(indus, nox)    
plot(lstat, medv)   
plot(lstat, rm)     
```

It seems that there is no relationship between zn and rm. There is a positive relationship between indus and nox. There is a negative relationship between lstat and medv. There is a negative relationship between lstat and medv.

#### (c)
```{r}
cor(crim, Boston)
```

Since the correlation of which is greater than 0.5 is dis and tax. There is positive relationship between crim and dis, tax.

#### (d)
```{r}
summary(crim)
```

The maximum of crime rate is  88.97%. And the median is 0.2565%. So there are suburbs have particularly high crime rate.
```{r}
summary(tax) 
```

The maximum of tax rate is 7.11%. The median is 3.3% and Q3 is 6.66%. So the max of tax rate is not so high.
```{r}
summary(ptratio)
```

The maximum of pupil-teacher ratios is 22 and the median is 19.05. So there is not any suburbs which have particularly high pupil-teacher ratios.

#### (e)
```{r}
sum(chas)
```

#### (f)
```{r}
median(ptratio)
```

#### (g)
```{r}
which.min(medv)  
Boston[which.min(medv),]
for (i in 1:13){
  print(paste("summary of", names(Boston)[i]))
  print(summary(Boston[,i]))
}
```

It is higher than 75% of the overall ranges of crim. It is the lowest of overall ranges of zn. It is 75% of overall ranges of indus. It is higher than 75% of the overall ranges of nox. It is lower than 25% of the overall ranges of rm. It is the max of the overall ranges of age. It is lower than 25% of the overall ranges of dis. It is the 75% of the overall ranges of rad. It is the 75% of the overall ranges of tax. It is the 75% of the overall ranges of ptratio. It is the max of the overall ranges of black. It is higher than 75% of the overall ranges of lstat.

#### (h)
```{r}
sum(rm>7)  
sum(rm>8) 
Boston[c(which((rm >8))),]
```

These 13 suburbs typically have a low crim rate, low(8 suburbs) or high zn, only 2 suburbs bound the river. nearly the same nox, close rm, high age except 2 suburb, only one high rad, nearly the same tax except one suburb, close pupil-teacher ratios and black, low lstat except one suburb and high medv except 3 suburbs.

### chapte 3 Problem 8
#### (a)
```{r}
library(ISLR)
data(Auto)
attach(Auto)
lm.mh <- lm(mpg~ horsepower)
summary(lm.mh) 
```

(1). There is a relationship between the predictor and the response since p-value is very very small. (2). The relationship is very strong since p-value is close to 0. (3). The relationship is negative since the estimate coefficient is negative.
```{r}
predict(lm.mh, data.frame(horsepower= 98), interval = "confidence")
predict(lm.mh, data.frame(horsepower= 98), interval = "prediction")
```

(4). fit: 24.46708, 95% confidence: (23.97308, 24.96108), 95% prediction: (14.8094, 34.12476)

#### (b)
```{r}
plot(horsepower, mpg)
abline(lm.mh, lwd = 2, col = "red")
```

#### (c)
```{r}
par(mfrow = c(2,2))
plot(lm.mh)
```

The plot of residuals-fitted values shows that there is non-linear relationship. The plot of Normal Q-Q shows that it is not a normal distirbution. The plot of residuals-leverage shows than there are a few outliers and a few high leverage points.

### chapter 3 Problem 15
#### (a)
```{r}
coeff_sim <- c()   # coeffient of simple linear model
for (i in 2:14){
  print(paste("crim~",names(Boston)[i]))
  lm.fit <- lm(crim~ Boston[,i])
  print(summary(lm.fit))
  plot(Boston[,i], crim, xlab = names(Boston)[i])  
  abline(lm.fit, lwd = 2, col = "red")
  # plot linear regression line for each model
  coeff_sim[i-1] <- coef(lm.fit)[2]  # extract coefficients
}
```

The p-value of all variables except chas are smaller than 0.001, which means except chas, all variables have relationship with crim.

#### (b)
```{r}
lm.all <- lm(crim~., data = Boston)
summary(lm.all)
```

For predictors zn, nox, dis, rad, black, lstat, medv, we can reject the null hypothesis, since their p-value is very small.

#### (c)
```{r}
coeff_mul <- coef(lm.all)[-1]
plot(coeff_sim, coeff_mul)
```

The simple coefficient and the multiple coefficient of nox is totally different. The others are very close.

#### (d)
```{r}
for (i in 2:13){
  print(paste("crim ~",names(Boston)[i]))
  lm.fit <- lm(crim~ Boston[,i]+I(Boston[,i]^2)+I(Boston[,i]^3))
  print(summary(lm.fit))
}
```

There is evidence of non-linear association between response and predictor zn, indus, nox, age, dis, ptratio, lstat.